---
title: "Economic Effects on Marriage"
author: "Brandon Huff"
date: "Revised on 2019-09-17"
output: 
  pdf_document:
    fig_height: 2
    fig_width: 3.5
    df_print: kable
---
```{r, setup, include=FALSE}
# Load necessary packages
require(mosaic)
require(kableExtra)
require(forcats)
require(lmtest)
require(emmeans)
require(Stat2Data)



# Set default chunk and display options
knitr::opts_chunk$set(
  tidy = F,     # display code as typed
  size = "small",   # slightly smaller font for code
  comment = "\t",   # tab plain R output
  fig.align = "center",   # center figures
  warning = F, message = F)   # suppress warnings & messages in knitted doc   
theme_set(theme_classic())   # greyscale ggplot them
options(scipen = 1) # don't be so quick to use scientific notation
```




# Variables of Interest
Sex (SEX)
Age (AGE)
Marital Status (MAR)
Year Last Married (MARHYP)
Divorced in the past 12 months (MARHD)
Family Income in the past 12 months (FINCP)
Region (REGION.x)
Family Income (FamIncome)
Personal Income (PersonalIncome)
Family presence and age of related children (FPARC)
Number of related children in household (NRC)
Work status of householder or spouse in family households (WORKSTAT)
Same-sex married couple households (SSMC)
# Data wrangling


## Rename Data
```{r}
# read in data
acs230 <- read.file("acs230_3k.csv")
# wrangle!!
acs230 <- rename(acs230, 
                 MaritalStatus = MAR,
                 Age = AGEP,
                 Sex = SEX,
                 Region = REGION.x,
                 Race = RAC1P,
                 YearMarried = MARHYP,
                 FamIncome = FINCP,
                 PersonalIncome = PINCP)

```
## Mutate Data
```{r}
# mutate
acs230 <- mutate(acs230, 
                 Race = fct_recode(as.factor(Race),
                           "White alone"= "1", 
                           "Black or AA alone" = "2",
                           "Asian alone" = "6",
                           "Two or More Races" = "9"),
                 Race = fct_collapse(as.factor(Race),
                                     "Other" = 
                                       c("3", "4", "5", "8", "7")),
                 MaritalStatus = fct_recode(as.factor(MaritalStatus),
                                    "Married" = "1",
                                    "Widowed" = "2",
                                    "Divorced" = "3",
                                    "Separated" = "4",
                                    "NeverMarried" = "5"),
                 Sex = fct_recode(as.factor(Sex),
                                  "Male" = "1",
                                  "Female" = "2"),
                 FPARC = fct_recode(as.factor(FPARC),
                                    "5 years only" = "1",
                                    "5-17 years only" = "2",
                                    "5 years and 17 years" = "3",
                                    "No children" = "4"),
                 BinaryIncome = ifelse(PersonalIncome<100000,
                                       FALSE, TRUE),
                 BinaryMarriage = ifelse(MaritalStatus=="Divorced",
                                         TRUE, FALSE),
                 SSMC = fct_recode(as.factor(SSMC),
                                   "No" = "0",
                                   "Yes" = "1",
                                   "All couples" = "2"),
                 LogPI = log(PersonalIncome),
                 AgeGrp = fct_collapse(as.factor(Age), 
                               "00-09" = c("00", "01", "02", "03", "04", "05",
                                         "06", "07", "08", "09"),
                               "10-19" = c("10", "11", "12", "13", "14", "15",
                                         "16", "17", "18", "19"),
                               "20-29" = c("20", "21", "22", "23", "24", "25",
                                         "26", "27", "28", "29"),
                               "30-09" = c("30", "31", "32", "33", "34", "35",
                                         "36", "37", "38", "39"),
                               "40-49" = c("40", "41", "42", "43", "44", "45",
                                         "46", "47", "48", "49"),
                               "50-59" = c("50", "51", "52", "53", "54", "55",
                                         "56", "57", "58", "59"),
                               "60-69" = c("60", "61", "62", "63", "64", "65",
                                         "66", "67", "68", "69"),
                               "70-79" = c("70", "71", "72", "73", "74", "75",
                                         "76", "77", "78", "79"),
                               "80-89" = c("80", "81", "82", "83", "84", "85",
                                         "86", "87", "88", "89"),
                               "90-99" = c("90", "91", "92", "93", "94", "95",
                                         "96", "97", "98", "99")))
  

```
## Filter
```{r}
adults <- filter(acs230, Age >= 18, PersonalIncome >0,)
data2 <- adults%>%filter(is.na(FPARC)==FALSE)
adults2 <- filter(adults, Age >=30 )
```

# Data exploration

Explore distributions and associations graphically and numerically.
```{r}
#View(adults)
# numeric summaires
group_by(acs230, Race) %>% summarize(meanAge = mean(Age))
```
```{r}
#age distribution
gf_histogram( ~ Age, data = adults2)
#number in each Age Group
group_by(adults2, AgeGrp)%>%
  summarize(Count = n())
group_by(adults2, Region)%>%
  summarize(Count = n())
group_by(adults2, Sex)%>%
  summarize(Count = n())
#Personal Income Distribution
gf_histogram( ~ PersonalIncome, data = adults2)
#Personal Income(log) distribution
gf_histogram( ~ LogPI, data = adults2)
#Marital Status distribution
gf_bar( ~ MaritalStatus, data = adults2, stat="count")
#Race distribution
gf_bar( ~ Race, data = adults2, stat="count")
#Number in each Race
group_by(adults2, Race)%>%
  summarize(RaceCount = n())
#FPARC distribution
group_by(adults2, FPARC)%>%
  summarize(FPARCCount = n())
#FPARC distribution
group_by(adults2, SSMC)%>%
  summarize(SSMCCount = n())
#Boxplot of each Marital Satus distribution of Personal Income
gf_boxplot(PersonalIncome ~ MaritalStatus, data = adults2)
#Marital Status table
group_by(adults2, MaritalStatus)%>%
  summarize(meanFamIncome = mean(FamIncome, na.rm = TRUE),
            meanPersonalIncome = mean(PersonalIncome, na.rm = TRUE),
            MaritalStatusCount = n(),
            meanAge = mean(Age))
#Personal Income Summary
summarize(adults2, mean = mean(PersonalIncome), standard = sd(PersonalIncome, na.rm = TRUE))
#NRC Summary
summarize(adults2, mean = mean(NRC, na.rm = TRUE), standard = sd(NRC, na.rm = TRUE))
#Year Married Summary
summarize(adults2, mean = mean(YearMarried, na.rm = TRUE), standard = sd(YearMarried, na.rm = TRUE))

```
```{r}
```


# Data analysis
## Multiple Regression Model

Perform analyses.
```{r}
lm01 <- lm(LogPI ~ MaritalStatus + Age + Sex + Region + Race + YearMarried
           + MARHD + FPARC +  NRC , data = adults2)
msummary(lm01)
anova(lm01)
```
```{r}
#drop Age
lm02 <- lm(LogPI ~ MaritalStatus + Sex + NRC + Region + Race + YearMarried
           + MARHD + FPARC + SSMC,data = adults2)
msummary(lm02)
anova(lm02)
```
```{r}
#drop MARHD variable
lm03 <- lm(LogPI ~ MaritalStatus + Region + Race + YearMarried
           + FPARC + Sex + NRC,data = adults2)
msummary(lm03)
anova(lm03)
```
```{r}
#drop Race variable
lm04 <- lm(LogPI ~ SSMC + MaritalStatus + Region + YearMarried
           + FPARC + NRC + Sex, data = adults2)
msummary(lm04)
anova(lm04)
```
```{r}
#drop Region
lm05 <- lm(LogPI ~ MaritalStatus + FPARC + NRC 
           + YearMarried + Sex + Age, data = adults2)
msummary(lm05)
anova(lm05)
```
```{r}
confint(lm05, 'Age')
```
```{r}
confint(lm05, 'YearMarried')
```
```{r}
confint(lm05, 'NRC')
```
```{r}
confint(lm05, 'SexFemale')
```

```{r}
confint(lm05, 'MaritalStatusWidowed')
confint(lm05, 'MaritalStatusDivorced')
confint(lm05, 'MaritalStatusSeparated')
```
```{r}
confint(lm05, 'FPARC5-17 years only')
confint(lm05, 'FPARC5 years and 17 years')
confint(lm05, 'FPARCNo childreny')
```

```{r}
mplot(lm05, c(1,2))
```
## Logistic Regression Model
```{r}
#BinaryMarriage distribution
group_by(adults2, BinaryMarriage)%>%
  summarize(Count = n())
```

```{r}
logmodel0 <- glm(BinaryMarriage ~ LogPI + Sex + Age
                + (LogPI * Sex) + (LogPI * Age), 
                data = adults2, family = binomial(logit))
msummary(logmodel0)
anova(logmodel0)
```
```{r}
#drop PersonalIcome and Age interaction term
logmodel <- glm(BinaryMarriage ~ LogPI + Sex + Age
                + (LogPI * Sex), 
                data = adults2, family = binomial(logit))
msummary(logmodel)
anova(logmodel)
```
```{r}
#drop Age 
logmodel1 <- glm(BinaryMarriage ~ PersonalIncome + Sex
                + (PersonalIncome * Sex), 
                data = adults2, family = binomial(logit))
msummary(logmodel1)
anova(logmodel1)
```
```{r}
#Empirical Logit Plot
emplogitplot1(BinaryMarriage ~ PersonalIncome + Sex + 
                (PersonalIncome * Sex), data = adults2, ngroups = 20)
```

## PostHoc Analysis
```{r}
PHmod1 <- lm(data = adults2, LogPI ~ Age + MaritalStatus + FPARC + NRC 
           + YearMarried + Sex) 
car::Anova(PHmod1, type = "III")
```

```{r}
# Pairwise comparisons of MaritalStatus using Tukey's adjustment
(pwMaritalStatus <- emmeans(PHmod1, specs = pairwise ~ MaritalStatus))
```
```{r}
 # Corresponding confidence intervals
confint(pwMaritalStatus)
```
```{r}
# Pairwise comparisons of FPARC using Tukey's adjustment
(pwFPARC <- emmeans(PHmod1, specs = pairwise ~ FPARC))
```
```{r}
 # Corresponding confidence intervals
confint(pwFPARC)
```
```{r}
# Pairwise comparisons of Sex using Tukey's adjustment
(pwSex <- emmeans(PHmod1, specs = pairwise ~ Sex))
```
```{r}
 # Corresponding confidence intervals
confint(pwSex)
```
```{r}
PHmod2 <- glm(data = adults2, BinaryMarriage ~ PersonalIncome + Sex
                + (PersonalIncome * Sex)) 
car::Anova(PHmod2, type = "III")
```

```{r}
confint(logmodel1, 'PersonalIncome')

```


